Attack Graph Generation for Micro-service Architecture

### 2018-09-18 paper_own_docker/chapters/background.tex
Todos in the background and minor changes in method and evaluation done

We first start by introducing the concept of microservices, their benefits and security implications in Subsection \ref{chap:microservices}. Afterward in Subsection \ref{chap:vulnerability_scanners}, we look into vulnerability scanners as tools that generate vulnerabilities for a single host . In the end in Subsection \ref{chap:attack_graphs}, we introduce and formally describe attack graphs as methods to diagnose security weaknesses of a given system composed of multiple hosts.

>>>>

We first start by introducing the concept of microservices, their benefits and security implications in Subsection \ref{chap:microservices}. Afterward in Subsection \ref{chap:vulnerability_scanners}, we look into vulnerability scanners as tools that generate vulnerabilities for a single host. At the end in Subsection \ref{chap:attack_graphs}, we introduce and formally describe attack graphs as methods to diagnose security weaknesses of a given system composed of multiple hosts.

### 2018-09-18 paper_own_docker/chapters/background_attack_graphs.tex
TODOs
Todos in the background and minor changes in method and evaluation done

The definition of attack graphs may vary but it is essentially a directed graph that consists of nodes and edges with various representations. In this subsection, we first look at a few examples of how others define attacks graphs and at the end present the model of an attack graph that we use in this work.

@@

Ou et al. introduce the notion of a logical attack graph. A logical attack graph is a bipartite directed graph that consists of two kinds of nodes: fact nodes and derivation nodes. Each fact node is labeled with a logical statement in the form of a predicate applied to its arguments, while each derivation node is labeled with an interaction rule that is used for the derivation step. The edges in the graph represent the "depends on" relation \cite{ou2006scalable}.

>>>>

Attack graphs are a popular way of examining network security weaknesses. They help security analysts to carefully analyze a given system and detect its vulnerable parts. The definition of attack graphs may vary but it is essentially a directed graph that consists of nodes and edges with various representations. In this subsection, we first look at a few examples of how others define attacks graphs and at the end present the model of an attack graph that we use in this work.

@@

Ou et al. introduce the notion of a logical attack graph. A logical attack graph is a bipartite directed graph that consists of two kinds of nodes: fact nodes and derivation nodes. Each fact node is labeled with a logical statement in the form of a predicate applied to its arguments, while each derivation node is labeled with an interaction rule that is used for the derivation step. The edges in the graph represent a "depends on" relation \cite{ou2006scalable}.

### 2018-09-18 paper_own_docker/chapters/background_microservices.tex
Todos in the background and minor changes in method and evaluation done

As real-world software increases in size, there is an ever growing need to decompose it into an organized structure to promote scaling, reuse and readability. A software application whose modules cannot be executed independently is called a monolith. Monoliths are characterized by tight coupling, vertical scaling and strong dependence. Service Oriented Architecture (SOA) addresses these issues by restructuring its elements into components that provide services which are used by other entities through a networking protocol \cite{papazoglou2003service}. However, in a typical SOA, the services are monolithic which gives rise to microservices in order to provide an even more fine-grained task separation \cite{ahmadvand2016requirements}.
In the microservices paradigm, multiple services are split into very basic units which are task oriented. According to Dragoni et al. a microservice is a cohesive, independent process interacting via messages. These microservices constitute a distributed architecture called a microservice architecture \cite{dragoni2017microservices}. Microservice architectures provide us with the advantage of having more heterogeneous technologies, cheaper scaling, resilience, organizational alignment, and composability among other benefits \cite{newman2015building}. However, they add an additional complexity and have a wider attack surface as the need of many services to communicate with each other and third-party software increases \cite{combe2016docker, dragoni2017microservices}.

>>>>

As real-world software increases in size, there is an ever growing need to decompose it into an organized structure to promote scaling, reuse and readability. A software application whose modules cannot be executed independently is called a monolith. Monoliths are characterized by tight coupling, vertical scaling and strong dependence. Service Oriented Architecture (SOA) addresses these issues by restructuring its elements into components that provide services which are used by other entities through a networking protocol \cite{papazoglou2003service}. However, in a typical SOA, the services are monolithic which gives rise to the concept of microservices in order to provide an even more fine-grained task separation \cite{ahmadvand2016requirements}. The novel term "microservices" was first introduced in 2011 at an architectural workshop in order to bridge a common term for the explorations of multiple researchers \cite{dragoni2017microservices, microservicesfrowler}. In the microservices paradigm, multiple services are split into very basic units which are task oriented. According to Dragoni et al. a microservice is a cohesive, independent process interacting via messages. These microservices constitute a distributed architecture called a microservice architecture \cite{dragoni2017microservices}. Microservice architectures provide us with the advantage of having more heterogeneous technologies, cheaper scaling, resilience, organizational alignment, and composability among other benefits \cite{newman2015building}. However, they add an additional complexity and have a wider attack surface as the need of many services to communicate with each other and third-party software increases \cite{combe2016docker, dragoni2017microservices}. While microservices are an architectural principle, container technology has emerged in cloud computing to provide a lightweight virtualization mechanism. This technology enables microservices to be packaged and orchestrated through the Cloud \cite{pahl2016microservices}.

### 2018-09-18 paper_own_docker/chapters/evaluation.tex
Todos in the background and minor changes in method and evaluation done

\begin{table*}
	\begin{center}
		\begin{tabular}{ p{20mm}p{35mm}p{35mm}p{10mm}cp{35mm} } 
			\hline
			Name & Description & Technology stack & No. Containers & No. vuln. & Github link \\\hline

			Netflix OSS & Combination of containers provided from Netflix. & Spring Cloud, Netflix Ribbon, Spring Cloud Netflix, Netflix's Eureka & 10 & 4111 & \url{https://github.com/Oreste-Luci/netflix-oss-example} \\

			Atsea Sample Shop App & An example online store application. & Spring Boot, React, NGINX, PostgreSQL & 4 & 120 & \url{https://github.com/dockersamples/atsea-sample-shop-app} \\

			JavaEE demo & An application for browsing movies along with other related functions. & Java EE application, React, Tomcat EE & 2 & 149 & \url{https://github.com/dockersamples/javaee-demo} \\

			PHPMailer and Samba & An artificial example created from two separate containers. We use an augmented version for the scalability tests. & PHPMailer(email creation and transfer class for PHP), Samba(SMB/CIFS networking protocol) & 2 & 548 &  \url{https://github.com/opsxcq/exploit-CVE-2016-10033}
			\url{https://github.com/opsxcq/exploit-CVE-2017-7494} \\

			\hline
		\end{tabular}
	\end{center}

	\caption{Microservice architecture examples analyzed by the attack graph generator.}
	\label{table_technologies}

\end{table*}

@@

\subsection{Use Cases}
\label{chap:heterogenious_systems}
Modern microservice architectures use an abundance of different technologies, number of containers, various connectivity and number of vulnerabilities. Therefore it is of immense importance to show that an attack graph generator works well in such heterogeneous scenarios. In order to do this, we tested our system on some real and slightly modified Github examples as described in Table \ref{table_technologies} Our intention was to find and test examples that are publicly available for possible future comparison characterized by different system properties(topologies, technologies, vulnerabilities) and coming from different usage domains. We also had to take into account that an overwhelming majority of the examples publicly available are small with only one or few containers, which made this search challenging. The resulting examples are as follows: NetflixOSS \cite{netflixoss}, Atsea Sample Shop App \cite{atsea} and JavaEE demo \cite{java}. NetflixOSS is a microservice system provided by Netflix that is composed of 10 containers and uses Spring Cloud, Netflix Ribbon, Netflix Eureka etc. Atsea Sample Shop App is an e-commerce sample web application composed of 4 containers and that uses Spring Boot, React, NGINX and PostgreSQL. JavaEE demo is a sample application for browsing movies that is composed of only two containers and uses JavaEE, React and Tomcat EE. We ran the attack graph generator and verified the resulting attack graphs manually based on domain knowledge and under the assumption that the output from Clair \cite{clair}, NVD attack vectors \cite{booth2013national} and the pre- and postconditions from the work of Aksu et al. \cite{aksu2018automated} are correct. After running the attack graph generator, the attack graphs for the Atsea Sample Shop app and the JavaEE demo are small as expected with few nodes and edges. However, the structure of the resulting attack graphs in the Netflix case was quite linear. This linearity is because each container is connected to a few other containers to reduce unnecessary communication and increase encapsulation. Therefore based on this connectivity an attacker needs to perform multiple intermediate steps in order to reach the target container. All of the examples terminated, there are no directed edges from containers with higher privileges to lower privileges, no duplication of nodes and no reflexive edges, which is in line with the previously mentioned monotonicity property. Additionally, we noticed that the running time of our system for each of these examples was short, and additional scalability tests are needed. The Phpmailer \cite{phpmailer} and Samba \cite{samba} system is an artificial example that we use and extend in the following subsection to perform these scalability tests.
\textbf{TODO explain the result of correctness, what did you find out i.e., they are correct to our knowledge maybe show one example of an attack graph generated, also explain more the linearity thing}

@@

Ou et al. provide some more extended study where they test their system(MulVAL) on more examples. They mention that the asymptotic CPU time is between $O(n^2)$ and $O(n^3)$, where n is the number of nodes (hosts). The performance of the system for 1000 fully connected nodes takes more than 1000 seconds to execute \cite{ou2006scalable}. % The authors also provide an evaluation where he MulVAL clearly outperforms the Sheyner's system .

@@

The total time of the attack graph generation procedure for 1000 fully connected hosts (784 seconds) outperforms Ou's results (1000 seconds). In the Sheyners's extended example(4 hosts, 8 atomic attacks and multiple vulnerabilites) the attack graph took 2 hours to create. Our attack graph procedure even for the bigger number of hosts(1000) shows faster attack graph generation time. It however performs worse than Ingols, but that is attributed to the usage of MP attack graph which is different from ours. From the results we can see that Breath-first Search can be used to generate attack graphs for increasing number of services and denser connectivity in microservices architectures.

>>>>

@@

\subsection{Use Cases}
\label{chap:heterogenious_systems}

Modern microservice architectures use an abundance of different technologies,  number of containers, various connectivity and number of vulnerabilities. Therefore it is of immense importance to show that an attack graph generator works well in such heterogeneous scenarios. In order to do this, we tested our system on some real and slightly modified Github examples as described in Table \ref{table_technologies}. Our intention was to find and test examples that are publicly available for possible future comparison characterized by different system properties(topologies, technologies, vulnerabilities) and coming from different usage domains. We also had to take into account that an overwhelming majority of the examples publicly available are small with only one or a few containers, which made this search challenging. The resulting examples are as follows: NetflixOSS \cite{netflixoss}, Atsea Sample Shop App \cite{atsea} and JavaEE demo \cite{java}. NetflixOSS is a microservice system provided by Netflix that is composed of 10 containers and uses Spring Cloud, Netflix Ribbon, Netflix Eureka etc. Atsea Sample Shop App is an e-commerce sample web application composed of 4 containers and that uses Spring Boot, React, NGINX and PostgreSQL. JavaEE demo is a sample application for browsing movies that is composed of only two containers and uses JavaEE, React and Tomcat EE. We ran the attack graph generator and verified the resulting attack graphs of the small examples manually based on domain knowledge and under the assumption that the output from Clair \cite{clair}, NVD attack vectors \cite{booth2013national} and the pre- and postconditions from the work of Aksu et al. \cite{aksu2018automated} are correct. After running the attack graph generator, the attack graphs for the Atsea Sample Shop app and the JavaEE demo are small as expected with few nodes and edges. The structure of the resulting Netflix attack graph had a nearly linear structure in which each node is connected to a small number of other nodes that form a chain of attacks. This linearity is because each container is connected to a few other containers to reduce unnecessary communication and increase encapsulation. Therefore based on this connectivity an attacker needs to perform multiple intermediate steps in order to reach the target container. All of the examples terminated, there are no directed edges from containers with higher privileges to lower privileges, no duplication of nodes and no reflexive edges, which is in line with the previously mentioned monotonicity property. Additionally, we noticed that the running time of our system for each of these examples was short, and additional scalability tests are needed. The Phpmailer \cite{phpmailer} and Samba \cite{samba} system is an artificial example that we use and extend in the following subsection to perform these scalability tests.

@@

Ou et al. provide some more extended study where they test their system(MulVAL) on more examples. They mention that the asymptotic CPU time is between $O(n^2)$ and $O(n^3)$, where n is the number of nodes (hosts). The performance of the system for 1000 fully connected nodes takes more than 1000 seconds to execute \cite{ou2006scalable}. % The authors also provide an evaluation where he MulVAL clearly outperforms the Sheyner's system.

@@

The total time of the attack graph generation procedure for 1000 fully connected hosts (784 seconds) outperforms Ou's results (1000 seconds). In the Sheyners's extended example(4 hosts, 8 atomic attacks and multiple vulnerabilities) the attack graph took 2 hours to create. Our attack graph procedure even for the bigger number of hosts(1000) shows faster attack graph generation time. It, however, performs worse than the generator from Ingols et al., but that is attributed to the usage of MP attack graph which is different from ours. From the aforementioned results, we can see that Breath-first Search can be used to generate attack graphs for an increasing number of services and denser connectivity in microservices architectures.

### 2018-09-18 paper_own_docker/chapters/method.tex
Todos in the background and minor changes in method and evaluation done

Privileges play a central role in this attack graph generation. We model the privileges in a hierarchy. The privileges in ascending order are None, VOS(User), VOS(Admin), OS(User) and OS(Admin). VOS means that the privilege is exclusive to a virtual machine, while not affecting the host machine. However in our case, unlike hosts connected in a computer networks, these privileges are adapted to images and not virtual machines. On the other side, The keyword OS means that the host machine has been compromised. Since VOS are hosted on some machines, and their exploitation does not imply exploitation of the host machine, they are in the lower level of hierarchy \cite{aksu2018automated}. None means that no privilege is obtained, User that only a subset of user level privileges are available, while Admin grants control over the whole system.

@@

In order to show how the attack graph generation works in practice, we present a small example. The example is taken from the Netflix OSS Github repository. Netflix OSS example is a Spring Cloud based microservices architecture that enables Service Discovery (Eureka), Circuit Breaker (Hystrix), Intelligent Routing (Zuul) and Client Side Load Balancing (Ribbon) \cite{netflixoss, springcloudnetflix}. Displayed in Figure \ref{TopologyGraph} is a subset of the example topology. The topology consists of "Outside" node, "Docker daemon" node, Zuul, Eureka and other nodes. According to Netflix, Zuul is an edge service that provides dynamic routing, monitoring, resiliency, security, and more \cite{netflixzuul}, while Eureka is a REST (Representational State Transfer) based service that is primarily used in the AWS cloud for locating services for the purpose of load balancing and failover of middle-tier servers \cite{netflixeureka}. In Figure \ref{AttackGraph} we can see a part of the resulting attack graph. Parts of both graphs have been intentionally omitted to reduce complexity. An example path that an attacker would take could be to first attack the Zuul container by exploiting the CVE-2016-10249 vulnerability by crafting an image file, which triggers a heap-based buffer overflow\footnote{\url{https://nvd.nist.gov/vuln/detail/CVE-2016-10249}} and gain USER privilege.  Then with this USER privilege, it can exploit the CVE-2015-7554 vulnerability on the same container via crafted field data in an extension tag in a TIFF image\footnote{\url{https://nvd.nist.gov/vuln/detail/CVE-2015-7554}} to gain ADMIN privilege. Once the ADMIN privilege has been obtained on Zuul, the attacker can attack the Eureka container by exploiting CVE-2017-7600 via another crafted image\footnote{\url{https://nvd.nist.gov/vuln/detail/CVE-2017-7600}} and gain ADMIN privilege. It is important to note that this is not the only path that the attacker can take in order to have ADMIN privileges on Eureka. Another path would be to exploit the CVE-2018-1124 vulnerability via creating entries in procfs by starting processes, which could result in crashes or arbitrary code execution\footnote{\url{https://nvd.nist.gov/vuln/detail/CVE-2018-1124}}. This vulnerability can be exploited by having only USER privilege on Zuul to gain directly ADMIN privileges of the Eureka container. Our attack graph generator shows both paths since it is of an interest to see every possible route in which a container can be compromised.

@@

We developed this system to be used with a specific version of Docker and Docker-Compose. This modular structure could however be easily extended to other versions of Docker-Compose, vulnerability scanners and microservice architectures by replacing the Vulnerability and Topology Parsers with custom ones and making sure that their outputs conform to the input of the Attack Graph Parser.

\textbf{TODO} need to rewrite the previous paragraph \\
\subsubsection{Topology Parser}
\label{chap:topology_p}

In order to generate an attack graph of a given system, we require an arrangement of its components and connections described by a system topology. The topology of Docker containers can be described at either runtime or statically by using Docker Compose. In our case, since we are doing static attack graph analysis, we use Docker Compose as our main tool in the beginning. Docker Compose provides a docker-compose.yml file which is used for extraction of the topology of the system. However different versions of docker-compose.yml, use different syntax. For example, older versions use the deprecated keyword "link", while newer ones use exclusively "networks", to denote a connection between two containers. In this work, we use the keyword "networks" as an indicator that a connection between two containers exists.

@@

Another consideration that we take into account is the privileged access. Some containers require certain privileges over the docker daemon in order to function properly. For example a user may want to run some hardware (web-cam \cite{webcam}) or some applications that demand higher privilege levels (Libvirt \cite{libvirt}) from Docker. In Docker, this is usually done either by mounting the Docker socket or specifying the keyword "privileged" in the docker-compose.yml file. An attacker with access to these containers has also access to the Docker daemon. Once the attacker has access to the Docker daemon, he has potential access to the whole microservice system, since every container is controlled and hosted by the daemon.

@@

>>>>

Privileges play a central role in this attack graph generation. We model the privileges in a hierarchy. The privileges in ascending order are None, VOS(User), VOS(Admin), OS(User) and OS(Admin). VOS means that the privilege is exclusive to a virtual machine, while not affecting the host machine. However in our case, unlike hosts connected in a computer network, these privileges are adapted to images and not virtual machines. On the other side, The keyword OS means that the host machine has been compromised. Since VOS are hosted on some machines, and their exploitation does not imply exploitation of the host machine, they are in the lower level of hierarchy \cite{aksu2018automated}. None means that no privilege is obtained, User that only a subset of user level privileges are available, while Admin grants control over the whole system.

@@

In order to show how the attack graph generation works in practice, we present a small example. The example is taken from the Netflix OSS Github repository. Netflix OSS example is a Spring Cloud-based microservices architecture that enables Service Discovery (Eureka), Circuit Breaker (Hystrix), Intelligent Routing (Zuul) and Client Side Load Balancing (Ribbon) \cite{netflixoss, springcloudnetflix}. Displayed in Figure \ref{TopologyGraph} is a subset of the example topology. The topology consists of "Outside" node, "Docker daemon" node, Zuul, Eureka and other nodes. According to Netflix, Zuul is an edge service that provides dynamic routing, monitoring, resiliency, security, and more \cite{netflixzuul}, while Eureka is a REST (Representational State Transfer) based service that is primarily used in the AWS cloud for locating services for the purpose of load balancing and failover of middle-tier servers \cite{netflixeureka}. In Figure \ref{AttackGraph} we can see a part of the resulting attack graph. Parts of both graphs have been intentionally omitted to reduce complexity. An example path that an attacker would take could be to first attack the Zuul container by exploiting the CVE-2016-10249 vulnerability by crafting an image file, which triggers a heap-based buffer overflow\footnote{\url{https://nvd.nist.gov/vuln/detail/CVE-2016-10249}} and gain USER privilege.  Then with this USER privilege, it can exploit the CVE-2015-7554 vulnerability on the same container via crafted field data in an extension tag in a TIFF image\footnote{\url{https://nvd.nist.gov/vuln/detail/CVE-2015-7554}} to gain ADMIN privilege. Once the ADMIN privilege has been obtained on Zuul, the attacker can attack the Eureka container by exploiting CVE-2017-7600 via another crafted image\footnote{\url{https://nvd.nist.gov/vuln/detail/CVE-2017-7600}} and gain ADMIN privilege. It is important to note that this is not the only path that the attacker can take in order to have ADMIN privileges on Eureka. Another path would be to exploit the CVE-2018-1124 vulnerability via creating entries in procfs by starting processes, which could result in crashes or arbitrary code execution\footnote{\url{https://nvd.nist.gov/vuln/detail/CVE-2018-1124}}. This vulnerability can be exploited by having only USER privilege on Zuul to gain directly ADMIN privileges of the Eureka container. Our attack graph generator shows both paths since it is of an interest to see every possible route in which a container can be compromised.

@@

We developed this system to be used with a specific version of Docker and Docker-Compose. This modular structure could, however, be easily extended to other versions of Docker-Compose, vulnerability scanners and microservice architectures by replacing the Vulnerability and Topology Parsers with custom ones and making sure that their outputs conform to the input of the Attack Graph Parser.

\subsubsection{Topology Parser}
\label{chap:topology_p}

In order to generate an attack graph of a given system, we require an arrangement of its components and connections described by a system topology. The topology of Docker containers can be described at either runtime or statically by using Docker Compose. In our case, since we are doing static attack graph analysis, we use Docker Compose as our main tool. Docker Compose provides a docker-compose.yml file which is used for extraction of the topology of the system. However different versions of docker-compose.yml, use different syntax. For example, older versions use the deprecated keyword "link", while newer ones use exclusively "networks", to denote a connection between two containers. In this work, we use the keyword "networks" as an indicator that a connection between two containers exists.

@@

Another consideration that we take into account is the privileged access. Some containers require certain privileges over the docker daemon in order to function properly. For example, a user may want to run some hardware (web-cam \cite{webcam}) or some applications that demand higher privilege levels (Libvirt \cite{libvirt}) from Docker. In Docker, this is usually done either by mounting the Docker socket or specifying the keyword "privileged" in the docker-compose.yml file. An attacker with access to these containers has also access to the Docker daemon. Once the attacker has access to the Docker daemon, he has potential access to the whole microservice system, since every container is controlled and hosted by the daemon.

@@

\begin{table*}[t]
	\begin{center}
		\begin{tabular}{ p{20mm}p{35mm}p{35mm}p{10mm}cp{35mm} } 
			\hline
			Name & Description & Technology stack & No. Containers & No. vuln. & Github link \\\hline

			Netflix OSS & Combination of containers provided from Netflix. & Spring Cloud, Netflix Ribbon, Spring Cloud Netflix, Netflix's Eureka & 10 & 4111 & \url{https://github.com/Oreste-Luci/netflix-oss-example} \\

			Atsea Sample Shop App & An example online store application. & Spring Boot, React, NGINX, PostgreSQL & 4 & 120 & \url{https://github.com/dockersamples/atsea-sample-shop-app} \\

			JavaEE demo & An application for browsing movies along with other related functions. & Java EE application, React, Tomcat EE & 2 & 149 & \url{https://github.com/dockersamples/javaee-demo} \\

			PHPMailer and Samba & An artificial example created from two separate containers. We use an augmented version for the scalability tests. & PHPMailer(email creation and transfer class for PHP), Samba(SMB/CIFS networking protocol) & 2 & 548 &  \url{https://github.com/opsxcq/exploit-CVE-2016-10033}
			\url{https://github.com/opsxcq/exploit-CVE-2017-7494} \\

			\hline
		\end{tabular}
	\end{center}

	\caption{Microservice architecture examples analyzed by the attack graph generator.}
	\label{table_technologies}

\end{table*}
